---
title: "MB3 Pilot Analysis"
author: "George"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(lme4)
require(tidyverse)
require(here)
require(kableExtra)
require(sjPlot)

# for Bayesian regression
require(brms)
require(ggmcmc)
require(tidybayes)
require(ggdist)
require(modelr)
```

## Load Data

Load pilot data (from Julien Mayor's lab).

```{r load-data}
d1 <- read_csv(here("pilot/data/first_session_babies.csv")) %>%
  rename(sd_LT_incongruent_trials = sd_LT_congruent_trials_1) 
d2 <- read_csv(here("pilot/data/second_session_babies.csv")) %>%
  rename(sd_LT_incongruent_trials = sd_LT_congruent_trials_1,
         looking_time = `looking time`) 

cdat <- bind_rows(d1, d2) %>%
  filter(!is.na(id)) %>%
  mutate(labID = 1)
```

Loaded data from `r length(unique(cdat$id))` participants.
Their age distribution is shown below.

```{r}
cdat %>% group_by(id, age_months) %>% 
  summarise(n=n()) %>%
  group_by(age_months) %>%
  summarise(Nch=n()) %>% kable()
```


## Plot



```{r, fig.width=6, fig.height=4.5, caption="Log(looking time) by trial type and trial number with a simulated main effect (size=.3) of trial type, and bootstrapped 95% confidence intervals."}
# test data
dat <- cdat %>% select(1:14) # remaining columns are averages and notes

fam_dat <- dat %>% filter(is.na(test_trial)) %>% 
  select(-test_trial, -test_trial_audio, -test_trial_audio_length_ms, 
         -test_trial_type, -test_version)

test_dat <- dat %>% filter(!is.na(test_trial)) %>% 
  select(-fam_trial_audio, -fam_version) %>%
  separate(test_trial_type, sep='_', 
           into = c("test_trial_type_ch", "test_trial_type_num")) %>%
  mutate(trial_type = ifelse(test_trial_type_ch == fam_condition, "same", "different"),
         log_looking_time = log(looking_time))

dag <- test_dat %>% group_by(id, fam_condition, age_months, trial_type) %>%
  summarise(looking_time = mean(log_looking_time)) %>% 
  group_by(trial_type, age_months) %>%
  tidyboot::tidyboot_mean(looking_time) # quite slow..

pos = position_dodge(width=.2)
ggplot(dag, aes(x=age_months, y=mean, group=trial_type, color=trial_type)) + 
  geom_point(aes(y=mean, x=age_months), pos=pos) + 
  ylab("log(looking time)") + xlab("Age (months)") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), pos=pos) + 
  theme_bw() + geom_smooth(method="lm")
```


## Regression

Preregistered analysis, but without per-lab random effects and procedure fixed effects: 
Infants' log(looking time) (DV) ~ 1 + familiarization order (ABB vs ABA) * trial_type + age * trial_type (same rule vs different rule at test) + trial_num * trial_type + (trial_num*trial_type | subject) 


```{r}

m1 <- lmer(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             trial * trial_type + (trial * trial_type | id), data=test_dat)
# singular, pruning random effects

m2 <- lmer(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             trial * trial_type + (1 | id), data=test_dat)
tab_model(m2)
```

```{r}
plot_model(m2, sort.est = T) + theme_bw()
```


## Bayesian Regression

```{r, message=F, warning=F, cache=T}
bm1 <- brm(log_looking_time ~ 1 + 
             fam_condition * trial_type +
             age_months * trial_type +
             trial * trial_type + (1 | id), 
           data = test_dat, 
           warmup = 1000, iter = 2000, chains = 4, 
           inits = "random", cores = 4, seed=42) 

summary(bm1)
```

```{r, echo=F, eval=F}
get_variables(bm1)
bdat <- ggs(bm1)

bm1 %>% 
  spread_draws(b_Intercept, b_fam_conditionABB, b_trial_typesame, b_age_months, b_trial 
              b_fam_conditionABB:trial_typesame, b_trial_typesame:age_months,
              b_trial_typesame:trial) %>%
  mutate(mean = b_Intercept + b_fam_conditionABB * b_trial_typesame +
           b_age_months * b_trial_typesame + 
           b_trial * b_trial_typesame) %>%
  ggplot(aes(y = fam_condition, x = mean)) +
  stat_halfeye()
```


Generate posterior predictive values.

```{r}
# posterior predictive
test_dat %>%
  data_grid(id, fam_condition, trial_type, age_months, trial) %>%
  add_fitted_draws(bm1) %>%
  ggplot(aes(x = .value, y = trial_type)) +
  facet_grid(. ~ fam_condition) +
  stat_pointinterval(.width = c(.66, .95)) + 
  theme_bw()
```

