---
title: "MB3 Data validation"
author: "tbd"
date: "5/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data validation

The goal here is to run the models necessary for testing the hypotheses. 

Input: data.frames that result from pre-processing, format needs to be defined

Output:  data summaries & tables, data visualisations, table with results from hypothesis tests

```{r data validation}

```

## Full model specification

```{r full model}

full_model <- lmer(log_looking_time ~ 1 +
     familiarization_rule * trial_type +
     age * trial_type +
     experimental_method * trial_type +
     multilingual_exposure * trial_type +
     trial_num * trial_type +
     trial_num * age +
     repetition * trial_type +
     (trial_num * trial_type | subject) +
     (trial_type + test_order | lab),
     data = main_df)
     
```
## Pruning order:

* Change Subject level random slope from interaction to two main effects: trial_type and trial_num
* Lab level slope trial_type
* Lab level slope test_order
* Subject level slope trial_num
* Subject level slope trial_type
* Lab intercept
* Subject intercept

```{r pruning}

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
